%===============================================================================
% Scale, Symmetry, and the Renormalization Group: A Lie-Theoretic Unification
% A Pedagogical Introduction
%===============================================================================
% Tufte-style LaTeX Book
%===============================================================================

\documentclass[justified,notoc]{tufte-book}

%===============================================================================
% PACKAGES
%===============================================================================
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{mathtools}
\usepackage{slashed}  % For Feynman slash notation
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{bm}
\usepackage{mathrsfs}
\usepackage{bbold}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{arrows,decorations.pathmorphing,decorations.markings}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{booktabs}
\usepackage{lipsum}
\usepackage{units}
\usepackage{fancyvrb}
\usepackage{enumitem}
\usepackage[most]{tcolorbox}

%===============================================================================
% TYPOGRAPHY AND COLORS
%===============================================================================
\definecolor{darkblue}{RGB}{0,51,102}
\definecolor{darkgreen}{RGB}{0,102,51}
\definecolor{darkred}{RGB}{153,0,0}
\definecolor{lightgray}{RGB}{245,245,245}

\hypersetup{
    colorlinks=true,
    linkcolor=darkblue,
    citecolor=darkgreen,
    urlcolor=darkred
}

%===============================================================================
% CUSTOM COMMANDS
%===============================================================================
\newcommand{\dd}{\mathrm{d}}
\newcommand{\pp}{\partial}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\MM}{\mathcal{M}}
\newcommand{\LL}{\mathcal{L}}
\newcommand{\HH}{\mathcal{H}}
\newcommand{\OO}{\mathcal{O}}
\newcommand{\FF}{\mathcal{F}}
\newcommand{\TT}{\mathcal{T}}
\newcommand{\GG}{\mathcal{G}}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\sgn}{\mathrm{sgn}}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\const}{\mathrm{const}}
\newcommand{\vev}[1]{\langle #1 \rangle}
\newcommand{\Lie}{\mathfrak}
\newcommand{\ad}{\mathrm{ad}}
\newcommand{\Ad}{\mathrm{Ad}}
\newcommand{\id}{\mathrm{id}}
\newcommand{\Hom}{\mathrm{Hom}}
\newcommand{\End}{\mathrm{End}}
\newcommand{\Ric}{\mathrm{Ric}}
\newcommand{\rank}{\mathrm{rank}}

% Differential operators
\newcommand{\ddt}{\frac{\dd}{\dd t}}
\newcommand{\ddmu}{\frac{\dd}{\dd \mu}}
\newcommand{\ppt}{\frac{\pp}{\pp t}}
\newcommand{\ppx}{\frac{\pp}{\pp x}}

% Beta function styling
\newcommand{\betaf}{\beta}
\newcommand{\gammaf}{\gamma}

%===============================================================================
% THEOREM ENVIRONMENTS
%===============================================================================
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

%===============================================================================
% WORKED EXAMPLE BOXES
%===============================================================================
% Protected citation command for use inside tcolorbox environments
% This bypasses tufte's sidenote-based citation which conflicts with tcolorbox
\makeatletter
\newcommand{\boxcite}[1]{%
  \begingroup
  \let\@tufte@sidenote@citation\@gobble% Disable sidenote citation
  \citep{#1}%
  \endgroup
}
\makeatother

\newtcolorbox{workedbox}[1][]{
  enhanced,
  breakable,
  colback=lightgray,
  colframe=darkblue,
  boxrule=0.5pt,
  left=6pt, right=6pt, top=6pt, bottom=6pt,
  fonttitle=\bfseries\sffamily,
  title={#1},
  % Support for problem/solution separation
  bicolor,
  colbacklower=white,
  middle=2pt
}

%===============================================================================
% CHAPTER SUMMARY BOXES
%===============================================================================
\newtcolorbox{summarybox}{
  enhanced,
  breakable,
  colback=white,
  colframe=darkblue,
  boxrule=1pt,
  arc=6pt,
  left=10pt, right=10pt, top=10pt, bottom=10pt,
  shadow={2pt}{-2pt}{0pt}{black!15},
  title={\sffamily\bfseries\large Chapter Summary},
  coltitle=white,
  colbacktitle=darkblue,
  attach boxed title to top center={yshift=-3mm},
  boxed title style={arc=3pt, boxrule=0pt, left=6pt, right=6pt}
}

% Styled subsection headers for use inside summarybox
\newcommand{\summaryheader}[1]{%
  \par\vspace{6pt}\noindent{\sffamily\bfseries\color{darkblue}#1}\par\vspace{2pt}%
}

%===============================================================================
% REMARK BOXES (for conceptual asides)
%===============================================================================
\newtcolorbox{remarkbox}[1][Remark]{
  enhanced,
  breakable,
  colback=blue!5!white,
  colframe=blue!50!black,
  boxrule=0.5pt,
  arc=4pt,
  left=8pt, right=8pt, top=8pt, bottom=8pt,
  fonttitle=\bfseries\sffamily\small,
  coltitle=blue!50!black,
  colbacktitle=blue!10!white,
  title={#1},
  attach boxed title to top left={yshift=-2mm, xshift=4mm},
  boxed title style={arc=2pt, boxrule=0.5pt, colframe=blue!50!black}
}

%===============================================================================
% SOLUTION BOXES
%===============================================================================
\newtcolorbox{solutionbox}[1][Solution]{
  enhanced,
  breakable,
  colback=white,
  colframe=darkgreen,
  boxrule=0.5pt,
  arc=4pt,
  left=8pt, right=8pt, top=8pt, bottom=8pt,
  fonttitle=\bfseries\sffamily\small,
  coltitle=darkgreen,
  colbacktitle=white,
  title={#1},
  attach boxed title to top left={yshift=-2mm, xshift=4mm},
  boxed title style={arc=2pt, boxrule=0.5pt, colframe=darkgreen}
}

%===============================================================================
% MARGIN NOTE STYLING
%===============================================================================
\setlength{\marginparsep}{1em}
\setlength{\marginparwidth}{2in}

%===============================================================================
% DOCUMENT INFO
%===============================================================================
\title{Scale, Symmetry, and the\\Renormalization Group}
\author{A Unified Treatment}
\publisher{From Dynamical Systems to Quantum Field Theory}
\date{\today}

%===============================================================================
% BEGIN DOCUMENT
%===============================================================================
\begin{document}

\frontmatter

%-------------------------------------------------------------------------------
% TITLE PAGE
%-------------------------------------------------------------------------------
\maketitle

%-------------------------------------------------------------------------------
% COPYRIGHT/DEDICATION
%-------------------------------------------------------------------------------
\newpage
\begin{fullwidth}
~\vfill
\thispagestyle{empty}
\setlength{\parindent}{0pt}
\setlength{\parskip}{\baselineskip}
Copyright \copyright\ \the\year

\par\smallcaps{github.com/muhammadhasyim/rg\_ideas}

\par\textit{First printing, \today}
\end{fullwidth}

%-------------------------------------------------------------------------------
% TABLE OF CONTENTS
%-------------------------------------------------------------------------------
\tableofcontents
\listoffigures
\listoftables

%-------------------------------------------------------------------------------
% PREFACE
%-------------------------------------------------------------------------------
\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}

This book is the culmination of years spent exploring the surprising connections between renormalization group ideas across seemingly disparate fields. My fascination began while studying Goldenfeld's \textit{Lectures on Phase Transitions and the Renormalization Group}, a text remarkable not only for its treatment of Wilsonian renormalization in statistical mechanics and field theory, but also for its exposition of Barenblatt's work on self-similar solutions in nonlinear porous media flow. That a single mathematical framework could describe both critical phenomena in magnets and the spreading of groundwater through rock struck me as deeply significant.

\marginnote{Goldenfeld's book remains an essential reference for anyone seeking to understand RG beyond the confines of a single discipline.}

This discovery became the impetus for a broader investigation. I began tracing the renormalization group through fluid turbulence, chaotic dynamics, and quantum field theory, finding in each case the same essential structure: a flow on a space of theories driven by changes in scale. What eventually emerged from these readings was an appreciation for the group-theoretic and geometric character of renormalization---the recognition that ``the renormalization group'' is quite literally a group (or more precisely, a semigroup) acting on a manifold of models, with the beta function as its infinitesimal generator.

Why, then, write another book on this subject when Goldenfeld's treatment already exists? The answer is that Goldenfeld's presentation, excellent as it is, was deliberately simplified for pedagogical purposes and does not develop the broader abstract framework that underlies the wide applicability of renormalization group methods. Much has happened in the decades since. Kunihiro and his students took the program to its fullest potential, applying RG techniques systematically to dynamical systems and unifying virtually all singular perturbation theories---work that proceeded in parallel to Goldenfeld's own contributions. Barenblatt laid the mathematical foundations for applications to partial differential equations through his theory of intermediate asymptotics. Fluid mechanicians developed RG approaches to turbulence. And a community of mathematical physicists has explored the Lie group structure and differential geometric foundations of the renormalization group with increasing rigor. The time has come to synthesize these developments into a single cohesive treatment.

The renormalization group stands as one of the most profound conceptual advances in twentieth-century physics. What began as a technical device for handling infinities in quantum electrodynamics has grown into a universal framework for understanding how physical systems behave across different scales. The same mathematical structure appears in statistical mechanics near phase transitions, in fluid turbulence, in chaotic dynamical systems, and throughout quantum field theory.

\marginnote{The unity of these phenomena under a single mathematical umbrella represents a triumph of physical abstraction comparable to the unification of electricity and magnetism.}

This book presents the renormalization group as an \textbf{exact geometric framework}. The space of all possible theories or models forms a manifold, and the renormalization group generates a flow on this manifold. Fixed points of the flow correspond to scale-invariant theories. The geometry of the manifold encodes deep physical information about how theories relate to one another. This framework exists \emph{independently} of how we compute with it.

\textbf{Perturbation theory} is the primary method for implementing the RG framework---computing beta functions, anomalous dimensions, and fixed point properties order by order in a small parameter. When perturbation theory is insufficient, \textbf{transseries methods} extend our computational reach to include non-perturbative physics.

Our pedagogical strategy reflects a clean separation between \textbf{structure} and \textbf{computation}:

\marginnote{Part I: the exact framework with simple examples. Part II: analytical methods. Part III: applications combining both.}

\textbf{Part I (Algebra and Geometry)} develops the exact geometric framework using \emph{simple, exactly solvable examples}. The anharmonic oscillator demonstrates running parameters and the resolution of secular terms. The amplitude equation (Hopf bifurcation normal form) provides a complete illustration of nontrivial fixed points, stability analysis, and universality---without requiring any loop calculations. The porous medium equation exhibits anomalous dimensions and second-kind self-similarity exactly. The $\phi^4$ theory introduces the full machinery while emphasizing geometric structure over perturbative details. A distinctive feature of our approach is that geometric structures---metrics, connections, curvature---are \emph{derived} from explicit calculations in these examples rather than postulated abstractly.

\textbf{Part II (Analysis)} develops the analytical methods for computing within the RG framework. Perturbation theory generically produces divergent series, but this divergence \emph{encodes} non-perturbative physics. The Borel transform, resummation, and resurgence theory provide tools for extracting physical predictions from divergent series. Transseries extend perturbation theory to include instanton and renormalon contributions.

\textbf{Part III (Applications)} applies the unified geometric-analytical framework to seven physical systems of increasing complexity: chaotic dynamics in the Lorenz equations, turbulence in fluids, fracture mechanics in solids (following Barenblatt's theory of intermediate asymptotics), phase transitions in the Ising and O(N) models, and quantum field theories including QED and the Hubbard model.

\section*{Prerequisites}

The reader should be familiar with undergraduate analysis and linear algebra, ordinary and partial differential equations, and elementary quantum mechanics. Some exposure to statistical mechanics and field theory will be helpful but is not strictly required. Each chapter is designed to be self-contained, developing the necessary mathematical background as the physics demands it.

\section*{Notation}

We adhere to the following notational conventions throughout the book.

\subsection*{Scale and Flow Parameters}
\begin{center}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{ll}
$\ell$ & Scale parameter (logarithm of energy or length scale) \\
$t$ & Time (when physical time is the flow parameter) \\
$\mu$ & Renormalization scale (energy units) \\
$\Lambda$ & UV cutoff (energy or momentum) \\
$\epsilon$ & Small expansion parameter
\end{tabular}
\end{center}

\subsection*{Parameter Space}
\begin{center}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{ll}
$g^i$ & Coupling constants (coordinates on theory space $\MM$) \\
$\beta^i(g)$ & Beta function (vector field generating RG flow) \\
$B^i{}_j$ & Stability matrix $\partial\beta^i/\partial g^j|_{g^*}$ at fixed point \\
$\Delta$ & Scaling dimension (eigenvalue of stability matrix) \\
$\gamma$ & Anomalous dimension (interaction correction to scaling) \\
$G_{ij}$ & Metric on theory space (Fisher/Zamolodchikov) \\
$\Gamma^i{}_{jk}$ & Connection on theory space
\end{tabular}
\end{center}

\subsection*{Perturbation Theory and Transseries}
\begin{center}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{ll}
$\tilde{f}(\epsilon)$ & Formal (divergent) power series \\
$\hat{f}_B(\zeta)$ & Borel transform of $\tilde{f}$ \\
$\mathcal{S}[\tilde{f}]$ & Borel sum (resummation) \\
$\mathcal{S}_\pm$ & Lateral resummations (above/below real axis) \\
$\sigma^n$ & Transseries parameters (for non-perturbative sectors) \\
$S_\omega$ & Stokes constant at singularity $\omega$ \\
$\Delta_\omega$ & Alien derivative probing singularity at $\omega$ \\
$\mathfrak{S}$ & Stokes automorphism
\end{tabular}
\end{center}

\subsection*{Standard Symbols}
\begin{center}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{ll}
$\dd$ & Differential ($\dd/\dd t$, $\dd x$) \\
$\pp$ & Partial derivative ($\pp/\pp x$) \\
$\RR, \CC, \ZZ, \NN$ & Real, complex, integer, natural numbers \\
$\MM$ & Theory space (parameter manifold) \\
$\vev{\cdot}$ & Expectation value $\langle \cdot \rangle$ \\
$O(\epsilon^n)$ & Terms of order $\epsilon^n$ and higher
\end{tabular}
\end{center}

\vspace{2\baselineskip}
\begin{flushright}
\textit{The Author}\\
\today
\end{flushright}

%===============================================================================
% MAIN MATTER
%===============================================================================
\mainmatter

%-------------------------------------------------------------------------------
% PROLOGUE (Before Parts)
%-------------------------------------------------------------------------------
\input{chapters/prologue}

%-------------------------------------------------------------------------------
% PART I: ALGEBRA AND GEOMETRY
%-------------------------------------------------------------------------------
\part{Algebra and Geometry}

\input{chapters/ch02_rg_geometry}
\input{chapters/ch03_fixed_points}

%-------------------------------------------------------------------------------
% PART II: ANALYSIS
%-------------------------------------------------------------------------------
\part{Analysis: Perturbation Theory and Resurgence}

\input{chapters/ch04_perturbation}
\input{chapters/ch05_algebra}
\input{chapters/ch06_resurgence}

%-------------------------------------------------------------------------------
% PART III: EXAMPLES AND APPLICATIONS
%-------------------------------------------------------------------------------
% \part{Examples and Applications}

% \input{chapters/ch07_lorenz}
% \input{chapters/ch08_fluids}
% \input{chapters/ch09_solids}
% \input{chapters/ch10_ising}
% \input{chapters/ch11_on_model}
% \input{chapters/ch12_hubbard}
% \input{chapters/ch13_qed}
% \input{chapters/ch14_unity}

%-------------------------------------------------------------------------------
% APPENDICES
%-------------------------------------------------------------------------------
\input{chapters/appendix_resurgence}

%===============================================================================
% BACK MATTER
%===============================================================================
\backmatter

%-------------------------------------------------------------------------------
% BIBLIOGRAPHY
%-------------------------------------------------------------------------------
\bibliographystyle{plainnat}
\bibliography{rg_book_refs}

\end{document}
